### 2016-07-06 coverletter.tex
updated coverletter for resubmission, approaches to get eps figure files if necessary, tidying up

\documentclass[12pt]{letter}

\usepackage[latin1]{inputenc}
\usepackage{url}
\usepackage[bookmarks]{hyperref}

\name{Gero Walter}
\signature{Gero Walter,\\ Louis Aslett,\\ Frank Coolen}
\address{Gero Walter\\ P.O. Box 513\\ 45600 MB Eindhoven\\ The Netherlands\\[2ex] Tel.\ +31 40 247 7316\\ \url{g.m.walter@tue.nl}}
%\telephone{}

\date{06.07.2016}
\location{Eindhoven}

\begin{document}
\longindentation=0pt
\begin{letter}{}
\opening{Dear Editors,}
please find enclosed our revised manuscript for the IJAR Special Issue on Bayesian Nonparametrics,
``Bayesian Nonparametric System Reliability using Sets of Priors''.
We would like to thank you for allowing us to resubmit a few days past the deadline.
We would also like to thank the two anonymous reviewers who provided valuable comments,
and hope that our revised manuscript can be considered for publication.
\closing{Best regards,}
\end{letter}
\end{document}